With Great Growth, Comes Greater Consumption: Energy Consumption of Countries Around the world in Relation to Their annual GDP growth

### 2024-04-22 paper/paper.qmd
Add tests, update paper,file clean up
update paper

---
title: "With great growth, comes greater Consumption: Energy consumption of countries around the world"
subtitle: "An analysis of economic growth and Energy consumption globally"
author: 
  - Rayan Awad Alim
thanks: "Code and data are available at: https://github.com/RayanAlim/Energy-Analysis"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

install_if_missing("readxl")
install_if_missing("here")
install_if_missing("dplyr")
install_if_missing("ggplot2")
install_if_missing("kableExtra")
install_if_missing("tidyr")
install_if_missing("caret")
```

@@

# Load 
library(readxl)
library(here)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(tidyr)
library(caret)

#Load Data 
#data <- read.csv(here("data", "analysis_data", "IDMC_processed.csv"))
```

@@

@@

The policy implications of these findings are significant, as they influence national energy strategies and their alignment with economic objectives. The transition towards sustainable energy sources is also a critical factor in this equation, with renewable energy adoption seen as a key element in sustaining long-term economic growth without the environmental degradation associated with fossil fuels [@Wei-2023].

@@

@@

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

@@

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

@@

3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- R language [@citeR]

>>>>

---
title: "With Great Growth, Comes Greater Consumption: Energy Consumption of Countries Around the world in Relation to Their annual GDP growth"
subtitle: "An analysis of economic growth and energy consumption globally between 1992 to 2022"
author: 
  - Rayan Awad Alim
thanks: "Code and data are available at: https://github.com/RayanAlim/Energy-Analysis"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

install_if_missing("readxl")
install_if_missing("here")
install_if_missing("dplyr")
install_if_missing("ggplot2")
install_if_missing("kableExtra")
install_if_missing("tidyr")
install_if_missing("caret")
install_if_missing("tinytex")

```

@@

# Load 
library(readxl)
library(here)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(tidyr)
library(caret)
library(tinytex)

#Load Data 
countries_energy_GDP_data <- read.csv(here("data", "analysis_data", "countries_energy_GDP_data.csv"))

income_group_energy_GDP_data <-
  read.csv(here("data", "analysis_data", "income_group_energy_GDP_data.csv"))
```

@@

## Background And Literature Review

@@

## Motivation
The policy implications of these findings are significant, as they influence national energy strategies and their alignment with economic objectives. The transition towards sustainable energy sources is also a critical factor in this equation, with renewable energy adoption seen as a key element in sustaining long-term economic growth without the environmental degradation associated with fossil fuels [@Wei-2023].

@@

The first dataset has the following variables shown in Table 1:
**Entity**
**Year**
**Total_Energy_Consumption_TWh**
**Average_GDP_Growth**

```{r, tab.cap = "Sample of countries Energy Consumption and GDP Growth Dataset", tab.width=5, echo=FALSE}
#| warning: false
#| message: false

countries_energy_GDP_data %>%
  select(Entity, Year, Total_Energy_Consumption_TWh, Average_GDP_Growth) %>%
  arrange(Entity) %>%
  slice(1:10) %>%
  kable(
    caption = "First ten rows of a dataset countries energy consumption and gdp growth",
    col.names = c("Country", "Year", "Energy Consumption (TWH)", "GDP Growth %"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

```
From the first dataset, we create a secoind one that is segmented by income level, it has has the following variables, a sample of 10 enteries is shown in Table 2:
**Standardized_Entity**
**Year**
**Total_Energy_Consumption_TWh**
**Average_GDP_Growth**

```{r, tab.cap = "Sample of countries grouped by income level and their Energy Consumption and GDP Growth Dataset from 1992-2022", tab.width=5, echo=FALSE}
#| warning: false
#| message: false

income_group_energy_GDP_data %>%
  select(Standardized_Entity, Year, Total_Energy_Consumption_TWh, Average_GDP_Growth) %>%
  arrange(Standardized_Entity) %>%
  slice(1:10) %>%
  kable(
    caption = "First ten rows of a dataset countries grouped by income level energy consumption and gdp growth",
    col.names = c("Income level group", "Year", "Energy Consumption (TWH)", "GDP Growth %"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

```

@@

Our models goal is to  to understand the dynamics between energy consumption and economic growth across various nations and years. Recognizing the critical role of energy in economic development, this analysis seeks to quantify the extent to which changes in energy consumption influence GDP growth.

@@

We make use of a linear regression model to investigate this relationship, formalized as follows:

$$
GDP\_growth_{it} = \beta_0 + \beta_1 \times Energy\_consumption_{it} + \epsilon_{it}
$$

Where:

$$GDP\_growth_{it}$$
- The GDP growth rate (%) for country \(i\) in year \(t\).
$$Energy\_consumption_{it}$$
- The primary energy consumption (in terawatt-hours, TWh) for country \(i\) in year \(t\).
 $$\beta_0$$
 - The intercept term, representing the baseline GDP growth when energy consumption is zero.
$$\beta_1$$
- The coefficient of energy consumption, reflecting the change in GDP growth rate associated with a one-unit increase in energy consumption.
$$\epsilon_{it}$$
- The error term, capturing all other factors affecting GDP growth not included in the model.

@@

- **Linear Assumption**: Preliminary data exploration suggested a potential linear relationship between the logarithm of energy consumption and GDP growth rates. This linear model provides a clear and interpretable framework to estimate the sensitivity of economic growth to changes in energy usage.

- **Economic Rationale**: Theoretically, energy is a fundamental input in production processes; thus, variations in energy supply are expected to directly impact economic output and growth. The linear regression model allows us to quantify this impact and provide insights valuable for policy formulation.

- **Statistical Considerations**: Linear regression is robust, widely understood, and provides comprehensive diagnostic tools to check for model assumptions. It serves as an excellent preliminary analytical tool, especially when the relationship does not involve complex interactions or non-linear dynamics.

- **Normality**: We assume the residuals of the regression, \( \epsilon_{it} \), are normally distributed, facilitating the use of classical statistical tests for hypothesis testing (e.g., t-tests for coefficients).
- **Independence**: Observations across years and countries are assumed to be independent of each other.
- **Homoscedasticity**: The error terms are assumed to have constant variance across different levels of energy consumption.
- **Linearity**: The relationship between GDP growth and energy consumption is presumed to be linear after appropriate transformations if required (e.g., logarithmic).
# Results

## Data Results

```{r, fig.cap = "Total GDP Growth of all copuntries Over Time 1992-2022", tab.width=5, echo=FALSE}
#| warning: false
#| message: false

ggplot(countries_energy_GDP_data, aes(x = Year, y = Average_GDP_Growth)) +
  geom_point(color="orange") +
  labs(title = "Average GDP Growth Over Time for High Income Countries",
       x = "Year",
       y = "GDP Growth (%)",
       caption = "Data source: Your data source") +
  theme_minimal()

```

```{r, fig.cap = "Total Energy Consumption Over Time 1992-2022", tab.width=5, echo=FALSE}
#| warning: false
#| message: false

# Define colors for each standardized entity
color_map <- c("High Income" = "blue", "Upper Middle Income" = "green",
               "Lower Middle Income" = "orange", "Low Income" = "red")

# Plotting Total Energy Consumption over time by income group
ggplot(income_group_energy_GDP_data, aes(x = Year, y = Total_Energy_Consumption_TWh, color = Standardized_Entity)) +
  geom_line(size = 1.2) +  # Using thicker lines for better visibility
  scale_color_manual(values = color_map) +
  labs(title = "Total Energy Consumption Over Time (1992-2022)",
       subtitle = "Data grouped by income level",
       x = "Year",
       y = "Total Energy Consumption (TWh)",
       color = "Income Group") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold", size = 14),
        plot.subtitle = element_text(face = "italic"),
        text = element_text(size = 12))
```

## Model Results

@@

3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The programming software R [@citeR].
